---
title: "COVID CASE STUDY"
author: "Mark Druss"
date: "21 June 2021"
output:
  word_document: default
  html_document: default
params:
       data_date:
              label: "Date of data:"
              value: "2021-06-21"
              input: date
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r Import_Packages}

pacman::p_load(
       rio,          # for importing data
       here,         # for locating files
       skimr,        # for reviewing the data
       janitor,      # for data cleaning  
       lubridate,    # for date cleaning  
       epikit,       # creating age categories
       gtsummary,    # creating tables  
       scales,       # percents in tables  
       viridis,      # color scales
       tsibble,      # epiweeks
       apyramid,     # age/sex pyramids
       flextable,    # for making pretty tables
       gghighlight,  # highlighting plot parts  
       ggExtra,      # special plotting functions
       tidylog,      # Records every data cleaning step
       tidyverse     # for data management and visualization
)
```


```{r Import_Data}

City_Pops <- import(here("learning_materials", "covid_case_study", "city_pops.xlsx"))

COVID_raw <- import(here("learning_materials", "covid_case_study", "covid_example_data.xlsx"))
```

```{r Exploratory Analysis, eval = FALSE}

#Take a look at the dataset
summary(COVID_raw)

```


```{r Clean_Data}

COVID <- COVID_raw %>% 
       
       # automatically clean column names
       clean_names() %>% 
       
       #Rename
       rename( #NEW                      #OLD
              Date_Report            = report_dt,
              Age                    = case_age,
              Gender                 = case_gender,
              Race                   = case_race,
              Ethnicity              = case_eth,
              Date_Symptom_Onset     = sym_startdt,
              Fever                  = sym_fever,
              Subject_fever          = sym_subjfever,
              Myalgia                = sym_myalgia,
              No_Taste_Smell         = sym_losstastesmell,
              Sore_Throat            = sym_sorethroat,
              Cough                  = sym_cough,
              Headache               = sym_headache,
              Symptoms_Resolved      = sym_resolved,
              Date_Sym_Resolve       = sym_resolveddt,
              Date_Hospital_admin    = hosp_admidt,
              Date_Hospital_discharge= hosp_dischdt,
              Covid_diagnosis        = covid_dx,
              Date_Died              = died_dt,
              Date_Positive_Sample   = pos_sampledt,
              Lat                    = latitude_jitt,
              Long                   = longitude_jitt) %>% 
       
       # convert date_onset to date class
       mutate(Date_Report = ymd(Date_Report),
              Date_Symptom_Onset      = ymd(Date_Symptom_Onset),
              Date_Sym_Resolve        = ymd(Date_Sym_Resolve),
              Date_Hospital_admin     = ymd(Date_Hospital_admin),
              Date_Hospital_discharge = ymd(Date_Hospital_discharge),
              Date_Positive_Sample    = ymd(Date_Positive_Sample),
              Date_Died               = ymd(Date_Died)) %>% 
                     
       #Filter for Confirmed cases
       filter(confirmed_case == "Yes") %>% 
       
       #Filter for before 21 June 2021
       filter(Date_Report <= ymd("2021-06-21")) %>% 
              
       # properly record missing values in many character columns
       mutate(across(.cols = where(is.character), .fns = na_if, "")) %>% 
              
       #Make an Age Group
       mutate(age.group = age_categories(Age,
                                         lower = 0,
                                         upper = 100,
                                         by = 10)) %>% 
       
       #Change Gender to NA
       mutate(Gender = na_if(Gender, "Unknown"))


```


# EXECUTIVE SUMMARY

The COVID outbreak continues although symptomatic cases are decreasing across all cities as we enter summer in Fulton County.

As of `r format(ymd(params$data_date), "%d %B, %Y")` there have been `r nrow(COVID)` confirmed cases reported and `r fmt_count(COVID, died == "Yes")`  deaths. 


# Person  

*Age Pyramid*  
```{r Age_Pyramid, fig.width=8}


# Age pyramid with ggplot modifications
age_pyramid(data = COVID,
            age_group = "age.group", #note that the column must be enclosed in quotation marks (" ")
            split_by = "Gender",  #note that the column must be enclosed in quotation marks (" ")
            proportional = TRUE,  #Change to a proportion
            na.rm = FALSE) +
       
       theme_minimal()+
       
       scale_fill_brewer(type = "qual", palette = 4)+
       
       labs(title = "Age and Gender of Confirmed COVID Cases",
            x = "Proportion of All Cases",
            y = "Age group",
            caption = "COVID outbreak DEMO",
            fill = "Gender")


```
<br>  

- As of `r format(ymd(params$data_date), format = "%B %d %Y")`, Fulton County has 
recorded `r nrow(COVID)` confirmed cases of COVID-19. Of these `r sum(COVID$Gender == "Male", na.rm = T)` are Male and `r sum(COVID$Gender == "Female", na.rm = T)` are Female. The average age is `r mean(COVID$Age, na.rm = T)` with a range of `r min(COVID$Age, na.rm = T)` to `r max(COVID$Age, na.rm = T)`.      

\pagebreak 

 
<br>
<br>
_Hospitalizations by Race_  
```{r}
# Compare via table Race and Outcome
#add p-value
#Select Columns

COVID %>% 
     select(Race, hospitalized) %>% 
     tbl_summary(by = hospitalized, percent = "row") %>% 
     add_p()

```

- Among all confirmed cases of COVID-19 in Fulton County, `r fmt_count(COVID, hospitalized == "Yes")`
required hospitalization.   

<br>
<br>  
_Deaths by Race_  
```{r}
COVID %>% 
     select(Race, died) %>% 
     tbl_summary(by = died, percent = "row")
```
<br>

- As of `r format(ymd(params$data_date), format = "%B %d %Y")`, Fulton County has 
recorded `r fmt_count(COVID, died == "Yes")` overall deaths.  

\pagebreak 

# Time  

```{r Weekly_Epi_Curve, fig.width=8}
weekly_breaks <- seq.Date(
      from = floor_date(min(COVID$Date_Report, na.rm=T),   "week", week_start = 1), # Monday before first case
      to   = ceiling_date(max(COVID$Date_Report, na.rm=T), "week", week_start = 1), # Monday after last case
      by   = "week")                                                                   # bins are 7-days 

# Run the plot, using the vector
ggplot(data = COVID, aes(x = Date_Symptom_Onset, fill = fct_lump_n(city, 5))) +
       geom_histogram(breaks = weekly_breaks, closed = "left") +
       
       scale_y_continuous(expand = c(0, 0)) +
       
       scale_x_date(
              date_breaks = "2 months",
              labels = scales::label_date_short()) +
       
       labs(x = "Date of Report", 
            y = "Incidence",
            fill = "City",
            caption = str_glue("Fake report for course credits\n{nrow(COVID)} cases shown by city of residence\nCases reported from {min(COVID$Date_Report)} to {max(COVID$Date_Report)} by Monday week")) +
       
       theme_bw()

#+       facet_wrap(~city)
```
  
-  A rapid decline of symtomatic cases is being reports as of `r format(ymd(params$data_date), format = "%B %d %Y")` for all cities in the county.  
<br>
<br>

# Place  

*Top 5 Cities reporting confirmed cases from `r min(COVID$Date_Report, na.rm=T)` to `r max(COVID$Date_Report, na.rm=T)`*   
     
```{r facets, fig.width=8}

# Run the plot, using the vector
ggplot(data = COVID, aes(x = Date_Symptom_Onset, fill = fct_lump_n(city, 5))) +
       geom_histogram(breaks = weekly_breaks, closed = "left") +
       
       scale_y_continuous(expand = c(0, 0)) +
       
       scale_x_date(
              date_breaks = "6 months",
              labels = scales::label_date_short()) +
       
       labs(x = "Date of Report", 
            y = "Incidence",
            fill = "City",
            caption = str_glue("Fake report for course credits\n{nrow(COVID)} cases shown by city of residence\nCases reported from {min(COVID$Date_Report)} to {max(COVID$Date_Report)} by Monday week")) +
       
       theme_bw()+       
       
       facet_wrap(~fct_lump_n(city, 5))


```

<br>
-  Atlanta is reporting the largest incidence rate decline in 2021 compared to the other Fulton County cities.  


\pagebreak 

*Cumulative Case Incidence Table by City*  
```{r}

COVID_Merged <- COVID %>% 
       
       group_by(city) %>% 
       
              summarise(n_cases    = n(),
                  recent_14d = sum(Date_Report < ymd(params$data_date) - 14),
                  n_confirmed = sum(confirmed_case == "Yes", na.rm = T),
                  n_deaths   = sum(died == "Yes", na.rm = T))%>% 
       
       left_join(City_Pops, by = "city") %>% 
       
       # add cululative case incidence  
       mutate(incidence_10k = round((n_cases / population) * 10000, digits = 1)) %>% 
       
       # remove extra column
       select(-population) %>% 
       
       # Decreasing order
       arrange(desc(incidence_10k)) 

```


```{r}

COVID_Merged %>% 
  qflextable %>%  
  set_header_labels(         
      city = "City", 
      n_cases = "Cases",
      n_confirmed = "Confirmed",
      n_deaths = "Deaths",
      total_cases = "Total COVID cases",                  
      recent_14d = "COVID Cases in last 14 days",
      incidence_10k = "Cumulative incidence per 10,000") 

```
# City Summary 

- As of `r format(ymd(params$data_date), format = "%B %d %Y")`, `r head(COVID_Merged[1,1], 1)` has 
recorded the highest incidence of confirmed COVID-19.  


\pagebreak 

# EXTRA


```{r Patient_Timeline, fig.width=9}
# Patient timeline visualization ------------------------------------------
# make data frame of only the first 20 hospitalizations in the epidemic
timelines <- COVID %>% 
     arrange(Date_Hospital_admin) %>%                 # sort dataset so that earliest are at the top
     tail(20) %>%                             # keep only the top 20 rows
     select(pid, starts_with("Date"))    # keep only certain columns 

# Pivot date columns longer
timelines_long <- timelines %>% 
     
     # pivot the dataset longer
     pivot_longer(
          cols = starts_with("Date"),
          names_to = "date_type",
          values_to = "date") %>% 
     
     # set the new column date_type as class factor, and define order for its values
     mutate(date_type = fct_relevel(date_type, "Date_Symptom_Onset", "Date_Report", "Date_Hospital_admin", "Date_Positive_Sample", "Date_Died"))

       
# plot the long dataset to show timelines for each patient
timelines_long %>% 
     ggplot(data = timelines_long,
            mapping = aes(
                 x = date,
                 y = pid,
                 color = date_type,
                 shape = date_type,
                 group = pid))+
     geom_point(size = 4)+
     geom_line()+
     theme_minimal()
```

\pagebreak 