---
title: "DEMO Situation Report"
subtitle: "Ebola Outbreak in Sierra Leone"
author: "Mark Druss"
output:
  word_document: default
date: "`r Sys.Date()`"
params:
  district: "West II"
  publish_date: "2014-12-01"
---


##  Purpose

I'm learning R because I seek:  

*  Analytical _flexibility_ and _ownership_.
*  Advanced _graphics_ and automated _reporting capability_.
*  To **support the movement towards free, open-source software**  

Today's date is `r Sys.Date()` and I am finishing the Intro to R Field Epidemiology class via AppliedEPI!


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

# 1. Load Packages

*  Using pacman() several packages were loaded for the analysis (e.g., lubridate, epikit, tidyverse)

```{r Load_Packages, eval = TRUE}

pacman::p_load(
     rio,          # for importing data
     here,         # for locating files
     skimr,        # for reviewing the data
     janitor,      # for data cleaning  
     lubridate,    # for date cleaning  
     epikit,       # creating age categories
     gtsummary,    # creating tables  
     scales,       # percents in tables  
     viridis,      # color scales
     tsibble,      # epiweeks
     apyramid,     # age/sex pyramids
     flextable,    # for making pretty tables
     gghighlight,  # highlighting plot parts  
     ggExtra,      # special plotting functions
     ggExtra,
     webshot,
     RColorBrewer,
     sf,           # for working with geospatial data
     ggspatial,    # for basemaps and north arrows
     raster,       # for spatial formatting  
     tidyverse     # for data management and visualization
)

```

# 2. Import Data  

*  Surveillance Linelist  
*  Hospital Linelists  
*  Lab Dataset  
*  Case Investigation Dataset  

```{r Import_Data, eval = TRUE}

# surveillance dataset
surv_raw <- import(here("data", "raw", "surveillance_linelist_20141201.csv"))

# hospital datasets
hosp_central  <- import(here("data", "raw", "hospitals", "20141201_hosp_central.csv"))
hosp_military <- import(here("data", "raw", "hospitals", "20141201_hosp_military.csv"))
hosp_other    <- import(here("data", "raw", "hospitals", "20141201_hosp_other.csv"))
hosp_port     <- import(here("data", "raw", "hospitals", "20141201_hosp_port.csv"))
hosp_smmh     <- import(here("data", "raw", "hospitals", "20141201_hosp_smmh.csv"))
hosp_missing  <- import(here("data", "raw", "hospitals", "20141201_hosp_missing.csv"))

# laboratory dataset
lab <- import(here("data", "raw", "lab_results_20141201.xlsx")) %>% 
     clean_names()

# Import lab data 
investigations <- import(here("data", "raw", "case_investigations_20141201.xlsx")) %>% 
     # remove unnecessary columns  
     select(-c(age, age_unit, gender))

# import linelist
combined <- import(here("data", "clean", "backup", "linelist_combined_20141201.rds"))

## import shapefile 
shapefile <- read_sf(here("data", "shp", "sle_adm3.shp"))


```


# 3. Clean the Surveillance Linelist

*  Cleaned the dataset using, clean_names(), rename(), select(), distinct(), mutate(), recode(), ifelse(), case_when(), filter().

```{r Clean_the_Surveillance_Linelist, eval = TRUE}

surv <- surv_raw %>% 
     
     # automatically clean column names
     clean_names() %>% 
     
     # manually clean column names   
     rename(
          date_onset = onset_date,
          date_report = date_of_report,
          district_res = adm3_name_res,
          district_det = adm3_name_det) %>%
     
     # remove unnecessary column
     select(-row_num) %>% 
     
     # de-duplicate rows  
     distinct() %>% 
     
     # convert date_onset to date class
     mutate(date_onset = mdy(date_onset)) %>% 
     mutate(date_report = mdy(date_report)) %>% 
     
     # convert age to numeric class
     mutate(age = as.numeric(age)) %>% 
     
     # convert "Unknown" gender to NA
     mutate(gender = na_if(gender, "Unknown")) %>% 
     
     # properly record missing values in many character columns
     mutate(across(.cols = where(is.character), .fns = na_if, "")) %>% 
     
     # re-code hospital column
     mutate(hospital = recode(hospital,
                              # for reference: OLD = NEW
                              "Mitilary Hospital"  = "Military Hospital",
                              "Port"               = "Port Hospital",
                              "Port Hopital"       = "Port Hospital",
                              "St. Mark's Maternity Hospital (SMMH)" = "SMMH")) %>%
     
     # recode gender
     mutate(gender = recode(gender,
                            "m" = "male",
                            "f" = "female")) %>% 
     
     # convert negative weight values to NA
     mutate(wt_kg = ifelse(wt_kg < 0, NA, wt_kg))  %>% 
     
     # create case definition
     mutate(case_def = case_when(
          lab_confirmed == TRUE             ~ "Confirmed",
          epilink == "yes" & fever == "yes" ~ "Suspect",
          TRUE                              ~ "To investigate")) %>% 
     
     # create age-in-years
     mutate(age_years = case_when(
          age_unit == "months" ~ age/12,   # if age is given in months
          age_unit == "years"  ~ age,      # if age is given in years
          is.na(age_unit)      ~ age)) %>% # if unit missing assume years, else NA
     
     # create age category column
     mutate(age_cat = age_categories(         # create new column
          age_years,                             # numeric column to make groups from
          lower = 0,
          upper = 70,
          by = 10)) %>% 
     
     # Make date-difference column  
     mutate(diff = date_report - date_onset) %>% 
     
     # create column marking TRUE if district of residence and detection differ
     mutate(moved = district_res != district_det) %>% 
     
     # create new column that prioritizes district of detection
     mutate(district = coalesce(district_det, district_res)) %>% 
     
     # remove suspect cases
     filter(case_def == "Confirmed") %>% 
     
     # re-arrange columns
     select(case_id, starts_with("date"), diff, gender, age, age_unit, age_years, age_cat, hospital, district, district_res, district_det, moved, everything())



```


# 4. Joins  

*  Combine a series of hospital datasets and join them to a surveillance linelist  
*  Join a _lab_ dataset  
*  Join a _investigation_ dataset

Then tidy up the new combined dataset and create *EPIWEEKS*

```{r Joins, eval = TRUE}

# Join other datasets to the surveillance linelist
hosp <- bind_rows(hosp_central, hosp_port, hosp_military, hosp_smmh, hosp_other, hosp_missing) %>% 
     # select specific columns from hosp, and re-name ID as case_ID
     select(
          case_id = ID,          # select and rename
          date_hospitalisation,  # select
          time_admission,        # select
          date_outcome,          # select
          outcome)               # select



# Join the two data frames with a left-join
combined <- left_join(surv, hosp, by = "case_id")


# Join the surveillance and hospital data frames with a left-join
combined <- left_join(combined, lab, by = "case_id")


# Join the two data frames with a left-join
combined <- left_join(combined, investigations, by = "case_id")


# Clean the new columns that have been joined to 'combined'
combined <- combined %>% 
     
     # convert all column names to lower case and remove spaces
     clean_names() %>% 
     
     # covert new columns to class date
     mutate(date_hospitalisation = mdy(date_hospitalisation),
            date_outcome         = mdy(date_outcome),
            date_infection       = ymd(date_infection)) %>% 
     
     # clean outcome and hospital missings
     mutate(outcome = na_if(outcome, ""),
            hospital = na_if(hospital, ""))

       #Create EPIWEEKS
combined <- combined %>% 
  mutate(week_onset      = yearweek(date_onset, week_start = 1), ## create week of onset variable  
         week_onset_date = as.Date(week_onset))                  ## create a date version 

```


# 5. Export the cleaned _combined_ dataset

*  Can export the dataset individually or part of a project.  

```{r Export, eval = TRUE}

# save the combined dataset
export(combined, here("data", "linelist_combined_20141201.rds"))


```

# 6. Practical Report Example

# Executive summary

The Ebola outbreak continues, although reported cases are decreasing in all districts following the surge of contract-tracing support provided by international partners.

As of `r format(ymd(params$data_date), "%d %B, %Y")` there have been `r nrow(combined)` confirmed cases reported.  


| Partner                                 | Arrived  | Activities                   |
|-----------------------------------------|----------|------------------------------|
| Doctors without Borders (MSF) - Belgium | 20 July  | Treatment centers            |
| World Health Organization (WHO)         | 22 July  | Surveillance and analytics   |
| Red Cross (ICRC)                        | 1 August | Community-based surveillance |

: Table of partners involved in the response


[Area intended to summarize trends and response activities]  



## Person

```{r, warning=F, message=F}
# Age pyramid with ggplot modifications
age_pyramid(
     data = combined,
     age_group = "age_cat",
     split_by = "gender",
     proportional = TRUE,
     show_midpoint = FALSE)+
     
  theme_minimal()+
     
  scale_fill_brewer(type = "qual", palette = "Dark2")+
  
  labs(title = str_glue("Age and Gender of Confirmed Cases"),
       x = "Age Group",
       y = "Proportion of All Cases",
       caption = str_glue("{nrow(combined)} cases were reported through {format(ymd(params$data_date), '%d %B, %Y')}. {fmt_count(combined, is.na(age))} cases are missing age and not shown. {fmt_count(combined, is.na(gender))} cases are missing gender and not shown."),
       fill = "Gender")


```

```{r}
#Age Pyramids====
age_pyramid(data = combined,
            age_group = "age_cat", #note that the column must be enclosed in quotation marks (" ")
            split_by = "gender")  #note that the column must be enclosed in quotation marks (" ")
```

```{r}
age_pyramid(data = combined,
            age_group = "age_cat", #note that the column must be enclosed in quotation marks (" ")
            split_by = "gender",
            proportional = T,
            na.rm = F) 
```


## Time

```{r, warning=F, message=F}

# define weekly breaks from Monday before first case

# Define and save the vector
ebola_weeks <- seq.Date(
  from = floor_date(min(combined$date_onset, na.rm=T), unit = "week", week_start = 1),
  to =   ceiling_date(max(combined$date_onset, na.rm=T), unit = "week", week_start = 1),
  by =   "week")


# Make epidemic curve
ggplot(data = combined)+
       geom_histogram(
          mapping = aes(x = date_onset),
          breaks = ebola_weeks,
          closed = "left")+
     
       # x-axis labels
       scale_x_date(
         expand            = c(0,0),           # remove excess x-axis space before and after case bars
         date_breaks       = "4 weeks",        # date labels and major vertical gridlines appear every 3 Monday weeks
         date_minor_breaks = "week",           # minor vertical lines appear every Monday week
         labels = scales::label_date_short())+ # automatically efficient date labels
       
       # y-axis
       scale_y_continuous(
         expand = c(0,0))+             # remove excess y-axis space below 0 (align histogram flush with x-axis)
       
       # aesthetic themes
       theme_minimal()+                # simplify plot background
       
       theme(
         plot.caption = element_text(hjust = 0,        # caption on left side
                                     face = "italic"), # caption in italics
         axis.title = element_text(face = "bold"))+    # axis titles in bold
       
       # labels including dynamic caption
       labs(
         title    = "Weekly Incidence of Cases (Weeks start Monday)",
         subtitle = "",
         x        = "Week of Symptom Onset",
         y        = "Weekly Incident Cases Reported",
         caption  = stringr::str_glue("n = {nrow(combined)} confirmed cases; Onsets range  {format(min(combined$date_onset, na.rm=T), format = '%a %d %b %Y')} to {format(max(combined$date_onset, na.rm=T), format = '%a %d %b %Y')} \n{nrow(combined %>% filter(is.na(date_onset)))} are missing date of onset and subsequently not shown"))
```

```{r, warning=F, message=F, fig.height= 7, fig.width=8}
# Make epidemic curve
ggplot(data = combined)+
       geom_histogram(
          mapping = aes(x = date_onset),
          breaks = ebola_weeks,
          closed = "left")+
     
       # x-axis labels
       scale_x_date(
         expand            = c(0,0),           # remove excess x-axis space before and after case bars
         date_breaks       = "4 weeks",        # date labels and major vertical gridlines appear every 3 Monday weeks
         date_minor_breaks = "week",           # minor vertical lines appear every Monday week
         labels = scales::label_date_short())+ # automatically efficient date labels
       
       # y-axis
       scale_y_continuous(
         expand = c(0,0))+             # remove excess y-axis space below 0 (align histogram flush with x-axis)
       
       # aesthetic themes
       theme_minimal()+                # simplify plot background
       
       facet_wrap(~ district, ncol = 2)+
     
       theme(
         plot.caption = element_text(hjust = 0,        # caption on left side
                                     face = "italic"), # caption in italics
         axis.title = element_text(face = "bold"))+    # axis titles in bold
       
       # labels including dynamic caption
       labs(
         title    = "Weekly Incidence of Cases (Monday weeks) by District",
         subtitle = "",
         x        = "Week of Symptom Onset",
         y        = "Weekly Incident Cases Reported",
         caption  = stringr::str_glue("n = {nrow(combined)} confirmed cases; Onsets range  {format(min(combined$date_onset, na.rm=T), format = '%a %d %b %Y')} to {format(max(combined$date_onset, na.rm=T), format = '%a %d %b %Y')}\n{nrow(combined %>% filter(is.na(date_onset)))} missing date of onset and not shown"))
```



## Place

### Summary by District  

```{r}
combined %>% 
     mutate(district = fct_explicit_na(district, na_level = "Missing")) %>% 
     tabyl(district) %>% 
     arrange(desc(n)) %>% 
     mutate(percent = scales::percent(percent, 1)) %>% 
     qflextable()
```


### Summary by Hospital  

```{r}
# table using {dplyr}
hospital_table <- combined %>% 
     mutate(hospital = fct_explicit_na(hospital, na_level = "Missing")) %>% 
     group_by(hospital) %>%                                     # get statistics for each hospital
     summarise(
          n_cases   = n(),                                         # number of rows (cases)
          max_onset = max(date_onset, na.rm = T),                  # latest onset date
          under5    = sum(age_years <= 5, na.rm = T),              # number of children under 5
          vomit_n   = sum(vomit == "yes", na.rm=T),                # number vomiting
          vomit_pct = percent(vomit_n / n_cases),                  # percent vomiting
          max_wt_male = max(wt_kg[gender == "male"], na.rm = T)) %>%     # max weight among men
     flextable::qflextable() %>% 
     set_header_labels(
          values = c(
               hospital = "Hospital",
               n_cases = "Total cases",
               max_onset = "Latest onset",
               under5 = "Under 5 years",
               vomit_n = "Vomit (n)",
               vomit_pct = "Vomit (%)",
               max_wt_male = "Max weight (males)"
          )
     )

# print
hospital_table
```

## District Spotlight  

Below is a spotlight on hospital admissions and outcomes for patients reported in
`r params$district`

```{r}
combined %>% 
  mutate(hospital = fct_explicit_na(hospital, na_level = "Missing")) %>% 
  filter(district == params$district) %>%   # filter data frame to the district in YAML
  drop_na(outcome) %>%                      # remove patients with missing outcome, for clarity
  tabyl(hospital, outcome) %>%              # begin the cross-tabulation    
  adorn_totals("both") %>%                  # add totals for both rows and columns
  qflextable() %>% 
  add_header_lines(str_glue("Outcomes in {params$district} District, by Hospital"))

```



# 7. Supplemental Example Plots

```{r, warning=F, message=F, fig.height= 10, fig.width=7}
ggplot(data = combined, mapping = aes(x = district_det, y = age_years, fill = district_det))+
       geom_jitter()+
       geom_boxplot(alpha = 0.5) +
       labs(title = "Ebola Outbreak DEMO Chart",
       subtitle = "Age by District Example",
       x = "District",
       y = "Age (years)",
       caption = "Among an Ebola outbreak, 2014",
       fill = "District")
```


```{r}

ggplot(data = combined, mapping = aes(x = district_det, fill = fct_rev(age_cat)))+
       geom_bar() +
       labs(title = "Ebola Outbreak DEMO Chart",
            subtitle = "Count of Cases by Ordered Age Group by Detected District Example",
            x = "District",
            y = "Count of Cases",
            caption = "Among an Ebola outbreak, 2014",
            fill = "Age Group")


```

```{r}

ggplot(data = combined, mapping = aes(x = district_det, fill = age_cat))+
       geom_bar(position = "dodge") +
       labs(title = "Ebola Outbreak DEMO Chart",
            subtitle = "Count of Cases by Age group by Detected District Example",
            x = "District",
            y = "Count of Cases",
            caption = "Among an Ebola outbreak, 2014",
            fill = "Age Group")


```

#Summarized data

```{r}
ggplot(data = combined, mapping = aes(x = age_years, y = ht_cm, color = age_cat))+
       geom_point()

```

```{r}
ggplot(
       data = combined,
       mapping = aes(
              x = district_det,
              fill = gender)) +
       geom_bar() +
       scale_fill_brewer(palette = "Pastel2",
                         na.value = "grey")
```


```{r}
#dynamic ages with gender
ggplot(data = combined, mapping = aes(x = age_years, y = ht_cm, color = gender))+
       geom_point()

```

```{r}

#using ggExtra

height_plot <- ggplot(data = combined, mapping = aes(x = age_years, y = ht_cm, color = gender))+
       geom_point()

ggMarginal(height_plot)

```


```{r}
ggMarginal(height_plot, groupFill = TRUE)
```

```{r}
#Stacked histo       
ggplot(data = combined, mapping = aes(x = age_years, fill = gender))+
       geom_histogram()
```

#FACETS

```{r}
ggplot(data = combined, mapping = aes(x = date_onset, fill = gender))+
       geom_histogram()+
       facet_wrap(~ district_det)
```

```{r}
ggplot(data = combined, mapping = aes(x = date_onset, fill = district_det))+
       geom_histogram()+
       facet_wrap(~ district_det)
```


```{r}
ggplot(data = combined, mapping = aes(x = date_onset, fill = district))+
       geom_histogram()+
       facet_wrap(~ district)+
       gghighlight()

#Add captio of 280 non finitie values and 28
```



```{r Create_Print_Plots, eval = TRUE}

# bar plot of case counts
ggplot(
     data = combined,
     mapping = aes(
          x = district,
          fill = gender)) +
     geom_bar() +
     scale_fill_viridis_d(na.value = "grey") +
     scale_y_continuous(breaks = seq(from = 0,
                                     to = 3000,
                                     by = 500),
                        expand = c(0,0)) +
     scale_x_discrete(expand = c(0,0))+
     coord_flip()


```



```{r}
# EPICurves ---------------------------------------------------------------

#Generic histo
ggplot(data = combined, mapping = aes(x = date_onset)) +
       geom_histogram(binwidth = 5) +
       labs(x = "Onset Date", y = "Incidence Count") +
       theme_minimal()
```

```{r}
#Defining the breaks, to associate with accurate counts within specific timeframes
seq.Date(from = ymd("2014-05-06"),
         to = ymd("2014-11-28"),
         by = "week")

#Then Merge to get 7 day intervals for binwidth
ggplot(data = combined, aes(x = date_onset))+
       geom_histogram(
              breaks = seq.Date(
                     from = ymd("2014-05-06"),
                     to = ymd("2014-11-28"),
                     by = "week")) +
       labs(x = "Date of onset", y = "Incidence") +
       theme_bw()
```


```{r}
#To make sure it starts on a certain day

# Monday BEFORE the earliest case
floor_date(ymd("2014-05-06"), unit = "week", week_start = 1)

# Monday AFTER the last case
ceiling_date(ymd("2014-11-28"), unit = "week", week_start = 1)

# Sequence of Mondays from before earliest case, to after latest case
#seq.Date(from = floor_date(ymd("2014-05-06"), unit = "week", week_start = 1),
         #to =   ceiling_date(ymd("2014-11-28"), unit = "week", week_start = 1),
         #by =   "week")

#To allow data to be updated DYNAMIC
seq.Date(from = floor_date(min(combined$date_onset, na.rm=T), unit = "week", week_start = 1),
         to =   ceiling_date(max(combined$date_onset, na.rm=T), unit = "week", week_start = 1),
         by =   "week")

#Then final merge save as an object

# Define and save the vector
ebola_weeks <- seq.Date(
       from = floor_date(min(combined$date_onset, na.rm=T), unit = "week", week_start = 1),
       to =   ceiling_date(max(combined$date_onset, na.rm=T), unit = "week", week_start = 1),
       by =   "week")
```



```{r}
#Add brackets to include at certain dates
# Define and save the vector
ebola_weeks <- seq.Date(
       from = floor_date(min(combined$date_onset, na.rm=T), unit = "week", week_start = 1),
       to =   ceiling_date(max(combined$date_onset, na.rm=T), unit = "week", week_start = 1),
       by =   "week")


# Run the plot, using the vector
ggplot(data = combined, aes(x = date_onset)) +
       geom_histogram(breaks = ebola_weeks, closed = "left") +
       labs(x = "Date of onset", y = "Incidence") +
       theme_bw()
```



```{r}

#To do by month
ebola_months <- seq.Date(
       from = floor_date(min(combined$date_onset, na.rm=T), unit = "month"),
       to =   ceiling_date(max(combined$date_onset, na.rm=T), unit = "month"),
       by =   "month")

#To do a facet wrap
ggplot(data = combined, aes(x = date_onset, fill = hospital)) +
       geom_histogram(breaks = ebola_months, closed = "left") +
       scale_x_date(
              date_breaks = "months",
              labels = scales::label_date_short())+
       labs(x = "Date of onset", y = "Incidence") +
       theme_bw()+
       facet_wrap(~hospital)
```



#Creating Proportions

```{r}

# weekly proportion of cases with more than 7 days delay between onset and report
delay_1wk <- combined %>%                                         # begin with surveillance linelist
     mutate(diff_1wk = as.numeric(diff) > 7) %>%              # create column that is TRUE is diff is greater than 7
     group_by(week = floor_date(date_report, "week")) %>%     # create column "week" and group by it  
     summarise(                                               # begin summarise command     
          cases = n(),                                           # number of cases in the week
          delayed = sum(diff_1wk == TRUE, na.rm=T),              # number of delayed cases in the week 
          delayed_pct = delayed / cases)                         # calculate proportion


# plot with dynamic labels
ggplot(data = delay_1wk, mapping = aes(x = week, y = delayed_pct))+
     geom_line(size = 2, color = "red")+
     labs(caption = str_glue(
          "n = {nrow(surv)}.\nReport produced on {Sys.Date()}\nData collected from {length(unique(surv$hospital))-2} major hospitals in the epidemic-affected area.\nLast reported case on {max(surv$date_report, na.rm = TRUE)}.\n{fmt_count(surv, is.na(date_report))} cases missing date of onset and not shown."))

```

Percents

```{r}

#To change y axis to percents

ggplot(data = delay_1wk, mapping = aes(x = week, y = delayed_pct))+
       geom_line(size = 2, color = "hotpink")+
       scale_y_continuous(labels = percent)+
       labs(caption = "n = 663.\nReport produced on 2022-04-02.Data collected from 5 major hospitals in the epidemic-affected area.\nLast reported case on 2014-12-21.\n7 cases missing date of onset.")


```


#Disease Progression via Case ID

```{r}
# Patient timeline visualization ------------------------------------------
# make data frame of only the first 5 cases in the epidemic
timelines <- combined %>% 
     arrange(date_onset) %>%                 # sort dataset so that earliest are at the top
     head(5) %>%                             # keep only the top 5 rows
     select(case_id, starts_with("date"))    # keep only certain columns 

# Pivot date columns longer
timelines_long <- timelines %>% 
     
     # pivot the dataset longer
     pivot_longer(
          cols = starts_with("date"),
          names_to = "date_type",
          values_to = "date") %>% 
     
     # set the new column date_type as class factor, and define order for its values
     mutate(date_type = fct_relevel(date_type, "date_infection", "date_onset", "date_report", "date_hospitalisation", "date_outcome"))

       
# plot the long dataset to show timelines for each patient
timelines_long %>% 
     ggplot(data = timelines_long,
            mapping = aes(
                 x = date,
                 y = case_id,
                 color = date_type,
                 shape = date_type,
                 group = case_id))+
     geom_point(size = 4)+
     geom_line()+
     theme_minimal()
```


# Highlighting top 3 Districts
```{r}
# Plot lumping together districts for plot (only show 3 most common and put others in "Other")  
ggplot(data = combined, 
       mapping = aes(
            x = date_onset,
            fill = fct_lump_n(district, 3)))+
     geom_histogram(binwidth = 7)+
     labs(fill = "District")
```


```{r}

# Epi plots using the combined dataset ------------------------------------
# make new data frame summarizing age, gender, and deaths
age_outcome_summary <- combined %>%
     drop_na(gender, age_cat) %>%         # remove any NA values
     group_by(age_cat, gender) %>%        # group rows by age-gender groups
     summarise(                           # begin creating new summary columns
          n = n(),
          n_death = sum(outcome == "Death",  # sum the number of rows where outcome is Death
                        na.rm = TRUE),
          pct_death = n_death / n)           # create proportion dead in the group
          
# make plot
ggplot(data = age_outcome_summary,            # use new data frame
       mapping = aes(x = gender,              # gender on x-axis
                     y = age_cat,             # age category on y-axis
                     fill = pct_death)) +     # fill (color) is shaded by the proportion dead
     geom_tile() +                               # display data as tiles  
     scale_fill_viridis(                         # adjust colors scale
          option = "B",                             # pick any option
          limits = c(0, 1)) +                       # set the limits to go from 0 to 1
     geom_text(                                  # add text over the tiles
          mapping = aes(
               label = scales::percent(                # show percents instead of proportions
                    pct_death,
                    accuracy = 0.1)))+
     labs(x = "Gender",                          # add labels
          y = "Age categories",
          fill = "Percent of\ncases fatal") +
     scale_fill_continuous(                      # adjust the values displayed in the legend
          labels = scales::percent)                 # to be percents (not proportions)


```

#Continuous Data

```{r}
#Example 3
ggplot(
       data = combined,
       mapping = aes(
              x = age_years,
              y = wt_kg,
              color = temp)) +
       geom_point() +
       scale_color_viridis_c(option = "inferno", na.value = "grey")
```

```{r}

ggplot(data = combined,
       mapping = aes(
              x = age_years,
              y = ht_cm,
              color = gender)) +
       geom_point(
              alpha = 0.7) +
       scale_color_brewer(
              palette = "Pastel2",
              na.value = "grey") +
       labs(
              title = "Height and age",
              subtitle = "All hospitals",
              x = "Age (years)",
              y = "Height (cm)",
              caption = "Fictional Ebola data",
              color = "Gender"
       ) +
       theme_classic(base_size = 16) +
       theme(
              legend.position = "bottom",                # move legend to bottom
              plot.title = element_text(color = "hotpink",   # title color
                                        size = 20,       # title font size
                                        face = "bold"),  # title typeface
              axis.title.y = element_text(angle = 0))    # rotate y axis title to be horizontal
```



#GIS Example

```{r}
#GIS
# Filter data -------------------------------------------------------------

## districts we are interested in 
districts <- combined %>% 
  distinct(admin3pcod) %>% 
  drop_na() %>% 
  pull() #pulls out the distinct values to be assigned to districts object

## filter shapefile for districts of interest 
shapefile <- shapefile %>% 
  filter(admin3Pcod %in% districts)


# Basic plot of the district shapes ---------------------------------------
## open up a ggplot
shape_plot <- ggplot() + 
  
  ## add the shapefile on top
  geom_sf(data = shapefile, 
          fill = NA,         # no fill
          colour = "black")  # black borders
# print
#shape_plot

```


```{r}

############# POINTS  ##########################################################

combined_sf <- combined %>% 
  drop_na(lat, lon) %>% 
  st_as_sf(                                               
    # define the coordinates based on lat/long variables
    coords = c("lon", "lat"),                             
    # set the coordinate reference system to WGS84
    crs = 4326,                                           
    # do not change string variables to factors 
    stringsAsFactors = FALSE                              
  )

# view the first 10 rows, first 5 columns, and the the geometry column
combined_sf$geometry

```


```{r}
## plot points on the district shapes
shape_plot + 
  geom_sf(data = combined_sf)+
  labs(title = "Case locations")
```


```{r}
## plot points on the district shapes, colored by outcome
shape_plot + 
     geom_sf(data = combined_sf,
             mapping = aes(color = fct_explicit_na(outcome)))+
     labs(color = "Outcome",
          title = "Case locations, by outcome") + 
  theme_minimal()

```


```{r}
############# CHOROPLETHS ####################################################

## get counts of cases by district
case_counts <- combined %>%
  count(admin3pcod, name = "counts")



## add case counts to the districts shapefile 
shapefile <- left_join(shapefile, case_counts, by = c("admin3Pcod" = "admin3pcod"))

# look at the districts shapefile and see the new column
View(shapefile)

## plot choropleth 
ggplot() + 
  ## add the shapefile on top
  geom_sf(data = shapefile, 
          # fill by case count
          aes(fill = counts),
          # black borders
          colour = "black")
```

